---
title: "README for the performance measurement part"
author: "Group 2"
date: "`December, 2nd, 2019`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document:
    df_print: paged
abstract: This Chapter is describes the evaluation of the best sampling and model combination, which is computed in the script 'PerformanceMeasurement.R'.
urlcolor: blue
---
## The goal of our performance measurement
After training a model it is important to check how well it performes. Since we not only trained one model but 140 in total (7 algos * 5 datasets * 4 positions), it is even more important to compare all the different models to see which one is the best. For our business case, this means that we want to see which combination of method and sampling we should take to predict which position, or if we even have a model, which does it better when we leave them together.

## How to compare the method/sampling combinations for all positions
After training the models with 10-fold cross-validation, we applied them on all the (unsampled) data from 2007 to 2013 to obtain the true positives, true negatives, false positives and false negatives. Since we used the cross-validation for training, this will not give us the training error, but the testing error we get with all the available data prior to the 2014 NFL Draft. At the end of every model-script, we save this information separately, to bring it all together in the script 'PerformanceMeasurement.R'.

The first step, is to bring them all into one dataframe, to use them easier. Then we make sure, that we used the same, unsampled data by computing the sums of TP, TN, FP, FN, for every method/sampling/position combination. In the CheckTibble we now see, that for every position the whole column contains always the same number, which means this is the case.

Then we calculate the ratio of correct classification, which is equal to:

$$\frac{\text{Correct Classifications}}{\text{All Classifications}} = \frac{TP + TN}{TP+TP+FP+FN}$$

The result is a table showing the percentage of correct classification for all the 140 model/sampling/position combination. This table is quite big, but it is still interesting to have a look at it, to see how well which combination performes.

```{r echo=FALSE}
load("../PerformanceMeasurement/BestModels.Rdata")
load("../PerformanceMeasurement/PerfMeasAllModels.Rdata")
```
```{r echo = FALSE}
knitr::kable(PerfMeasTibble, caption = "Ratio of correct classification", digits=4)
```

We can see these two tendencies in the models:

* Better performance with unsampled data, than with sampled data
* Better performance when we split the positions manually

## Our best models
With regards to the business case, we now need to decide which model/models we will use to predict the 2014 NFL Draft. Here you can see the best model/sampling combination for all positions.

```{r echo = FALSE}
knitr::kable(ResultTibble, caption = "The best model/sampling combinations by position", digits=4)
```

## Discussion
The great and sometimes perfect results of the random forest model show a high risk of a massive overfit. The other models with accuracies between 89% and 93% seem to be very good at the first sight. But we have to keep the reason for filtering out the players with <10 played games and sampling the data in mind. Here we applied the models to the unsampled but filtered data, which only contains 12.4% of drafted players. This means, that a model predicting "not drafted" for every player would not perform much worse, since it would still have an accuracy of 87.6%. 

Interpretations of models are always quite difficult, but the following thoughts are pretty likely to be true, according to the TP/TN/FP/FN. Our models are pretty good at predicting the likelihood of players being drafted that didn't perform very well in college football being low. It probably also does not too bad in predicting great players to be drafted, that nearly must be picked (and probably are picked early in the draft). But there is probably much room for improvement for all the players that still performed well in college, and might or might not be drafted.

We would like to close the circle to one of our fist lessons in machine learning, in which we were taught the following very high level formula for models (the right part). Y denotes the true outcome (in our case whether a player is drafted or not), f(X) is the true pattern that describes it, and $\epsilon$ is the noise, which appears to be random. With our models (the left part) we try to predict a $\hat{Y}$, which shall be as close as possible to the true Y.

$$\hat{f}(X) = \hat{Y} \approx Y = f(X) + \epsilon$$
Looking at this inequation we can think of three possibilities, why our models make so many mistakes:

* Our models $\hat{f_i}(X)$ do not include enough variables and/or are not sophisticated enough
* The data is not good enough
* The NFL draft contains a pretty large $\epsilon$ 



